\chapter{First Meeting Discussion}

\subsubsection{Work Package}
\begin{itemize}
\item Exercises
\item Problems
\item Tutorial
\item Appendix
\item Correction (Suggestion)
\end{itemize}

\subsubsection{Discussion}
\begin{itemize}
\item At the beginning, we first meet several times to agree on the style (format). For the next phase, I will discuss with PhD students to make sure the solutions are correct and elegant.
\item For the first edition, I will simply write down the exercise index without exercise statement. After finishing the main part, let's consider whether to state the exercises again in the solution set.
\item In the \LaTeX\ homework, commands of \LaTeX\ and \TeX\ are mixed. Maybe we should consider renewing the template.
\item Maybe we can add a Theorem collection or index as an appendix.
\item In Page 71, I think it's better to write $B$ instead of $B(\tau)$ because it is not the standard form of convolution, although they lead to the same result.
\item In Page 71, I think you mean ``In some cases this can be done directly from the \textbf{finite} series''.
\item In Page 74, maybe we should use symmetric instead of diagnoal?
\item Is Definitioin 5.1 rigorous?
\end{itemize}

\subsubsection{Questions}
\begin{itemize}
\item Am I going to have a position in the office?
\item You corrected typos with an additional document instead of fixing them in the lecture notes. Is there something wrong with the source file?
\item What is the difference between ``linear in $u$ and $x$'' and ``linear jointly in $u$ and $x$''?
\item When does the system diverge exponentially?
\item Is the script your work or a student's work?
\item What does $\{u_j\}_{j=1}^n \xrightarrow[]{A \in F^{m \times n}} \{v_i\}_{i=1}^m$ mean?
\end{itemize}